---
title: "IST8X Lecture 15 - Network Homework"
output:
  learnr::tutorial:
    progressive: false
    id: "network_homework"
    version: 0.1
runtime: shiny_prerendered
description: "Homework for the UC Davis DataLab IST8X course"
---

```{r setup, include=FALSE}
# packages
library(learnr)
library(statnet)
library(visNetwork)

# options
knitr::opts_chunk$set(echo = FALSE)

# load in data
toy_attributes = read.csv("./data/toy_attributes.csv", header = TRUE, stringsAsFactors = FALSE)
toy_edgelist = read.csv("./data/toy_edgelist.csv", header = TRUE, stringsAsFactors = FALSE)

# answers log location
.answers_log = "./YOUR_NAME_network_homework.answers"

# recorder function
my_recorder = function(tutorial_id, tutorial_version, user_id, event, data) {
  
  # write to log if answer
  if(!is.null(data$question)){
    
    # print metadata if student name
    if(data$question == "What is your name?"){
      cat("\n---------------------------------------------------------------\n",
          "Student Name: ", data$answer, "\n",
          "Time Started: ", as.character(Sys.time()), "\n",
          "System: ", Sys.info()["sysname"], "\n",
          "---------------------------------------------------------------\n",
          file = .answers_log, append = TRUE)
    } else {
      cat("\n", data$question, "\n", "Answer: ", data$answer, "\n", "Correct? ", data$correct, "\n", file = .answers_log, append = TRUE)
    }
  }
}

options(tutorial.event_recorder = my_recorder)

# network create
## sort your attributes frame alphabetically. Super important!
toy_attributes = toy_attributes[order(toy_attributes$id), ]

## make network!
toy_network = network(toy_edgelist, directed = FALSE)

# find what nodes are part of the largest component
toy_network%v%"lc" = component.largest(toy_network)
# delete those nodes that are not
## in the network
toy_network = delete.vertices(toy_network, which(toy_network%v%"lc" == FALSE))
## in our dataframes
toy_attributes = toy_attributes[toy_attributes$id %in% as.character(toy_network%v%"vertex.names"),]
toy_edgelist = toy_edgelist[which(toy_edgelist$to %in% toy_attributes$id | toy_edgelist$from %in% toy_attributes$id),]

# read in hw data
hw_attributes = read.csv("./data/hw_attributes.csv", header = TRUE, stringsAsFactors = FALSE)
hw_edgelist = read.csv("./data/hw_edgelist.csv", header = TRUE, stringsAsFactors = FALSE)

hw_net = network(hw_edgelist, directed = FALSE)
```

## Introduction

This is the interactive homework assignment for the UC Davis DataLab IST8X network lecture. It will present questions for you to answer along with spaces to save your responses. Make sure you hit submit or save for all questions!

You can re-try any problem in this assignment, or you can clear the assignment by clicking "Start Over" on the right.

When you are done, look in your working directory for `YOUR_NAME_network_homework.answers`. (Your working directory is `r getwd()`). This is what you will upload to Canvas. Please replace "YOUR_NAME" with your name.

Here is the first example of a question. Please enter your name here and hit "Submit Answer". Once you are done, hit "Next Topic" to advance to the next page.

```{r student_name}
question("What is your name?",
         answer(text = "", correct = TRUE),
         type = "learnr_text",
         incorrect = "Recorded (Ignore the red)!", allow_retry = TRUE, try_again_button = "Re-enter Name",
         options = list(placeholder = "John Smith"))
```

## Problem Set 1 - Network Data

This section will ask about network data.

For the following two questions, please select if the statement is TRUE or FALSE.

```{r net_data}
question("\nQ1.1 - An edgelist details what nodes in your network are connected.",
  answer("TRUE", correct = TRUE),
  answer("FALSE"),
  random_answer_order = TRUE, allow_retry = TRUE)
```

```{r att_list}
question("\nQ1.2 - An attributes table describes the edges between nodes (weights, frequency, etc.).",
  answer("TRUE", message = "The attributes table describes the *nodes* not the edges. Edge information is stored in the edgelist or adjacency matrix."),
  answer("FALSE", correct = TRUE),
  random_answer_order = TRUE, allow_retry = TRUE)
```

Use this adjacency matrix to answer the following questions.

```{r}
knitr::kable(data.frame("A" = c(1,0,1,1), "B" = c(0,1,0,0), "C" = c(1,0,1,0), "D" = c(1,0,0,1), row.names = c("A", "B", "C", "D")))
```

```{r weighted}
question("\nQ1.3 - If this adjacency matrix were turned into a network, would there be weighted edges?",
  answer("No", correct = TRUE),
  answer("Yes", message = "All the edges in this matrix are equal to 1, meaning they are imply binary yes/no rather than weighted."),
  random_answer_order = TRUE, allow_retry = TRUE)
```

```{r edge_guess}
question("\nQ1.4 - Which of the following edges exist in the network (Select all that apply)?",
  answer("A-B"),
  answer("A-C", correct = TRUE),
  answer("D-A", correct = TRUE),
  answer("C-B"),
  answer("B-D"),
  random_answer_order = TRUE, allow_retry = TRUE)
```

## Problem Set 2 - Coding and Measures

### Graph Level Properties

```{r}
# add pop-up tooltips with names
# visNetwork uses the "title" column to create pop-up boxes
toy_attributes$title = toy_attributes$id

# plot!
visNetwork(nodes = toy_attributes, edges = toy_edgelist) %>% visInteraction(zoomView = FALSE)
```

Using the network above, please answer the following questions.

```{r directed}
question("\nQ2.1 - Is this network directed or un-directed?",
              answer("Un-Directed", correct = TRUE),
              answer("Directed"),
              random_answer_order = TRUE)
```

```{r density_code, exercise=TRUE}
# replace the X with the proper function to find the density of the network
X(toy_network)
```

```{r density}
question_text("\nQ2.2 - What is the density of this network?",
              answer(text = as.character(round(gden(toy_network), digits = 3)), correct = TRUE),
              allow_retry = TRUE,
              placeholder = "Include 3 decimal places (e.g. 0.000)")
```

### Eigenvector Centrality

```{r}
toy_network%v%"evc" = evcent(toy_network, gmode = "graph", rescale = TRUE)

toy_attributes$title = toy_attributes$id

# reset all nodes to grey
toy_attributes$color = c("grey")

# replace min average geodesic with blue, max with red
toy_attributes$color[which(toy_network%v%"evc" == max(toy_network%v%"evc"))] = "red"
toy_attributes$color[which(toy_network%v%"evc" == min(toy_network%v%"evc"))] = "blue"

# plot
visNetwork(toy_attributes, toy_edgelist) %>% visInteraction(zoomView = FALSE)
```

```{r max_betweenness_code, exercise=TRUE}
# replace X with the proper function to find the eigenvector centrality of nodes in the network.
# Assign the values to the attributes dataframe.
toy_attributes$evc = X(toy_network, ...)

# View the dataframe
toy_attributes
```

```{r max_betweenness}
question_text("\nQ2.3 - What is the highest eigenvector centrality in the above network?",
              answer(text = as.character(round(max(evcent(toy_network, gmode = "graph", rescale = TRUE)), digits = 3)), correct = TRUE),
              allow_retry = TRUE,
              placeholder = "Include 3 decimal places (e.g. 0.000)")
```

<div id="max_betweenness_code-hint">
**Hint:** You can run code on the dataframe above! Use a function you learned previously on the betweenness column you made to find the maximum.
</div>

```{r max_betweenness_open}
textAreaInput("max_betweenness_open_text", "Q2.4 - What is the significance of a high eigenvector centrality score?")
actionButton("max_betweenness_open_save_button", label = "Save")
verbatimTextOutput("max_betweenness_open_saved")
```

```{r max_betweenness_open_back, context="server"}
max_betweenness_open_saved_text = eventReactive(input$max_betweenness_open_save_button, {
  cat("\nQ2.4 - What is the significance of a high betweenness centrality score?\n", input$max_betweenness_open_text, file = .answers_log, append = TRUE)
  paste0("Saved Answer: ", input$max_betweenness_open_text)
  })

output$max_betweenness_open_saved = renderText({max_betweenness_open_saved_text()})
```

## Problem Set 3 - Interpretation

```{r}
# add pop-up tooltips with names
# visNetwork uses the "title" column to create pop-up boxes
hw_attributes$title = hw_attributes$id

# plot!
visNetwork(nodes = hw_attributes, edges = hw_edgelist) %>% visInteraction(zoomView = FALSE)
```

Please use the above network to answer the following questions.

```{r bet_3}
textAreaInput("bet_3_text", "Q3.1 - Which node in this network do you think has the highest betweenness centrality score? Why? (Hover over the nodes to see names)")
actionButton("bet_3_save_button", label = "Save")
verbatimTextOutput("bet_3_saved")
```

```{r bet_3_back, context="server"}
bet_3_saved_text = eventReactive(input$bet_3_save_button, {
  cat("\nQ2.4 - Which node in this network do you think would have the highest betweenness centrality score? Why?\n", input$bet_3_text, file = .answers_log, append = TRUE)
  paste0("Saved Answer: ", input$bet_3_text)
  })

output$bet_3_saved = renderText({bet_3_saved_text()})
```





